Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-16 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
Fix section 4

In this section, we present an algorithm for \textsc{Online Multiclass
Classification with Bandit Feedback}, namely
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}.
Under the assumption that the examples are strongly linearly separable with a
margin, we give an upper bound on its number of mistakes. This is stated as
\autoref{theorem:strongly-separable-examples-mistake-upper-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

@@

The upper bound $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$ on the expected number of
mistakes of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} is
optimal upto a constant factor, as long as the number of classes $K$ is at most
$O((R/\gamma)^2)$. This lower bound is stated as
\autoref{theorem:strongly-separable-examples-mistake-lower-bound} below. The
proof of the theorem can be foudn in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

@@

\paragraph{Remark.} Irrespective of any conditions on $K$, $R$, and $\gamma$, a
trivial lower bound on the expected number of mistakes of any randomized
algorithm is $\frac{K-1}{2}$. To see this, note that the adversary can choose an
example $(R e_1, y)$, where $y$ is a label chosen uniformly from
$\cbr{1,2\ldots,K}$, and show this example $K-1$ times. The sequence of examples
trivially satisfies the strong linear separability condition, and the
$\frac{K-1}{2}$ expected mistake lower bound follows from Claim 2
of~\citet{Daniely-Helbertal-2013}. However, it is unclear if the trivial lower
bound is the best possible if $K$ is large, i.e., $\omega((R/\gamma)^2)$. We
leave this as an open problem.

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} can
be extended to nonlinear classification using
kernels~\cite{Scholkopf-Smola-2002, Shawe-Taylor-Cristianini-2004}. For each
class, as opposed to explicitly maintaining the predictor's weights for each
class, we maintain the set of examples whose scaled feature transformations are
added to the weights in Perceptron updates. We give a formal description of the
kernelized algorithm in Algorithm~\ref{algorithm:kernelized}.

The following theorem is a direct consequence of
Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound}.

\begin{theorem}[Mistake upper bound for kernelized algorithm]
\label{theorem:kernelized-upper-bound}
Let $k(\cdot, \cdot)$ be a kernel function with feature map $\phi$, $K$ be a positive
integer, $\gamma$ be a positive real number, $R$ be a non-negative real number. If
$L = ((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples in
$V \times \{1,2,\dots,K\}$ such that:
\begin{enumerate}
  \item $((\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T))$ is strongly linearly separable with margin $\gamma$,
  \item for all $t$ in $\cbr{1,\ldots,T}$, $k(x_t,x_t) \le R^2$,
\end{enumerate}
then the
expected number of mistakes
Algorithm~\ref{algorithm:kernelized}
makes is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}

>>>>

In this section, we consider the case when the examples are strongly linearly
separable. We present an algorithm for this setting
(Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})
and we give an upper bound on its number of mistakes, stated as
\autoref{theorem:strongly-separable-examples-mistake-upper-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

@@

The upper bound $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$ on the expected number of
mistakes of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} is
optimal upto a constant factor, as long as the number of classes $K$ is at most
$O((R/\gamma)^2)$. This lower bound is stated as
\autoref{theorem:strongly-separable-examples-mistake-lower-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

@@

\paragraph{Remark.} Irrespective of any conditions on $K$, $R$, and $\gamma$, a
trivial lower bound on the expected number of mistakes of any randomized
algorithm is $\frac{K-1}{2}$. To see this, note that the adversary can choose an
example $(R e_1, y)$, where $y$ is a label chosen uniformly from
$\cbr{1,2,\dots,K}$, and show this example $K$ times. The sequence of examples
trivially satisfies the strong linear separability condition, and the
$\frac{K-1}{2}$ expected mistake lower bound follows from Claim 2
of~\citet{Daniely-Helbertal-2013}. However, it is unclear if the trivial lower
bound is the best possible if $K$ is large, i.e., $\omega((R/\gamma)^2)$. We
leave this as an open problem.

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} can
be extended to nonlinear classification using
kernels~\cite{Scholkopf-Smola-2002, Shawe-Taylor-Cristianini-2004}. The formal
description of the kernelized algorithm is given in
Algorithm~\ref{algorithm:kernelized}. As opposed to explicitly maintaining the
weight vector for each class, the algorithm maintains the set of example-scalar
pairs corresponding to the updates of the non-kernelized algorithm. As a direct
consequence of
Theorem~\ref{theorem:strongly-separable-examples-mistake-upper-bound} we get a
mistake bound for the kernelized algorithm.

\begin{theorem}[Mistake upper bound for kernelized algorithm]
\label{theorem:kernelized-upper-bound}
Let $k(\cdot, \cdot)$ be a positive definite kernel function and let
$\phi(\cdot)$ be the associated feature map. Let $K$ be a positive integer,
$\gamma$ be a positive real number, $R$ be a non-negative real number. If $L =
((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples
such that:
\begin{enumerate}
  \item $((\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T))$ is strongly linearly separable with margin $\gamma$,
  \item for all $t$ in $\cbr{1,\ldots,T}$, $k(x_t,x_t) \le R^2$,
\end{enumerate}
then the expected number of mistakes Algorithm~\ref{algorithm:kernelized} makes
is at most $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$.
\end{theorem}
